Merging Datasets Through Deep learning

### 2018-08-28 paper2018/datasets.tex
cleaning up

\section{Datasets}
\label{datasets}
Our datasets were derived from Wikidata.  Specifically, we used the \textit{also known as} property from Wikidata to get alternate forms for the same entity name for people as well as companies.  For company names, we augmented the names and surface forms in Wikidata with data from the SEC \footnote{\url{https://www.sec.gov/dera/data/financial-statement-and-notes-data-set.html}}, which has former and more recent names for companies.  There were 213106 names for people from the specific dump we extracted, and 70946 names for companies.  The extracted files and the cleansing code are available on our repository \footnote{\url{https://github.com/yehudagale/fuzzyJoiner}}. The extracted files however contained noise that we cleaned up programmatically.  We describe the clean up rules for people and for companies separately because they were somewhat different.  In the case of people's names, we also augmented the data so the system could learn some common rules that define variants of a person's name.  This was not possible for company names.

@@

\subsection{Cleansing company names}
As with people's names, we removed any text in parentheses because it usually was a qualification (e.g., IBM (company)).  We restricted ourselves to unicode ISO-8859-1.  The SEC data had a lot of strange company names that could be described with the regex pattern T[0-9]+ or [0-9]+, and we dropped these.  We tried to ensure we included names that shared some subset of characters with the main name, ensuring we would not drop acronyms when possible.

@@

>>>>

\section{Benchmarks}
\label{datasets}
Our benchmarks were derived from Wikidata.  Specifically, we used the \textit{also known as} property from Wikidata to get alternate forms for the same entity name for people as well as companies.  For company names, we augmented the names and surface forms in Wikidata with data from the SEC \footnote{\url{https://www.sec.gov/dera/data/financial-statement-and-notes-data-set.html}}, which has former and more recent names for companies.  There were 213106 names for people from the specific dump we extracted, and 70946 names for companies.  The extracted files and the cleansing code are available on our repository. The extracted files however contained noise that we cleaned up programmatically.  We describe the clean up rules for people and for companies separately because they were somewhat different.  In the case of people's names, we also augmented the data so the system could learn some common rules that define variants of a person's name.  This was not possible for company names.

@@

\subsection{Cleansing company names}
As with people's names, we removed any text in parentheses because it usually was a qualification (e.g., IBM (company)).  We restricted ourselves to unicode ISO-8859-1.  The SEC data had a lot of strange company names that could be described with the regex pattern T[0-9]+ or [0-9]+, and we dropped these.  We tried to ensure we included names that shared some subset of characters with the main name, ensuring we would not drop acronyms when possible.  The check for acronyms tested if any of the initial letters of each name part occurred in the name.

@@

The data was then split with a 60-20-20 ratio to provide training, validation and test data respectively.  Each run was conducted with a different random split to ensure generalization of results.

### 2018-08-28 paper2018/evaluation.tex
cleaning up

Table\ref{Characteristics} quantifies the difficulty of the metric problem for the people and company data in Wikidata.  After conversion to character embeddings, we computed the 20 nearest neighbors for each anchor, and computed what percentage of positives were found within the top 20 (recall), the average distance of positives and negatives from the anchor and their standard deviations respectively.  To compare it with metric learning for face recognition, which has been studied before, we also provide the same metrics for Labeled Faces in the Wild based on their input vector embeddings.  The table quantifies the difficulty of the metric learning problem across different data types, with positive distances for each being greater than negative distances.  Baseline recall rates are also poor, which is important if one needs to establish that the neural network did in fact learn after training.

@@

Table\ref{Evaluation} shows the same metrics after training.  Because we compared across different losses, the results are catagorized by each loss function for each dataset we tested.  For each value reported in the table, we conducted two runs to ensure the results were stable across runs, and we report the mean in the table.

>>>>

Table \ref{Characteristics} quantifies the difficulty of the metric problem for the people and company data in Wikidata.  After conversion to character embeddings, we computed the 20 nearest neighbors for each anchor, and computed what percentage of positives were found within the top 20 (recall), the average distance of positives and negatives from the anchor and their standard deviations respectively.  To compare it with metric learning for face recognition, which has been studied before, we also provide the same metrics for Labeled Faces in the Wild based on their input vector embeddings.  The table quantifies the difficulty of the metric learning problem across different data types, with positive distances for each being greater than negative distances.  Baseline recall rates are also poor, which is important because it suggests that a different mapping function learnt by a neural network is in fact needed for these entity types, just as it is for faces.

@@

Table \ref{Evaluation} shows the same metrics after training.  Because we compared across different losses, the results are catagorized by each loss function for each dataset we tested.  For each value reported in the table, we conducted two runs to ensure the results were stable across runs, and we report the mean in the table.

### 2018-08-28 paper2018/introduction2.tex
cleaning up

Extensions in data management systems for handling `fuzzy' joins across different surface forms for the same entity typically use string similarity algorithms such as edit-distance, Jaro-Winkler and TF-IDF (e.g., \cite{Cohen2003}).  In applying fuzzy joins, a key problem is to avoid quadratic comparisons between pairs of strings drawn from the two columns.  A blocking strategy is used to reduce the number of pairs for fuzzy joins, such that only strings that share a common prefix, or suffix will be compared further.  String matching algorithms often do not work however, because transformations of entity names in perfectly valid ways can yield very different strings.  As shown in Table~\ref{table-example}, \textit{John Smith} is more similar to \textit{John Adams} than \textit{Adams, John}, but of course it is \textit{Adams, John} that is the valid alternate surface form for \textit{John Adams}.

More recently, data driven approaches have emerged as a powerful alternative to string matching techniques for fuzzy joins.  Data driven approaches mine patterns in the data to determine the `rules' for joining a given entity type.  One example of such an approach is illustrated in \cite{He:2015:SJS:2824032.2824036}, which determines which cell values should be joined based on whether those cell values co-occur on the same row across disparate tables in a very large corpus of data.  Another example of a data driven approach is work by \cite{auto-join-joining-tables-leveraging-transformations} where program synthesis techniques are used to learn the right set of transformations needed to perform the entity matching operation.

In this paper, we propose a novel approach to the problem of joining different surface representations of the same entity, inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a space partitioning algorithm to find the forms that can be potentially joined together.  Our approach for merging datasets is much more flexible than prior data driven approaches (e.g., \cite{He:2015:SJS:2824032.2824036}, \cite{auto-join-joining-tables-leveraging-transformations}) because it can be applied to any data provided one can create a mapping function such that joinable surface forms of the entity are closer in vector space than disparate forms.  For instance, in face recognition and person re-identification, there has been a significant research effort aimed at creating deep learning models that learn to map different surface forms of a face closer together in a vector space.  In our proposed approach, we can now join two columns if they have alternate faces for the same entity, whereas in prior work, those faces needed to either have been joined before in some existing dataset, or the transforms of a face need to be discoverable as a program, which is near impossible for faces.  Our approach for merging datasets is also efficient because it eliminates quadratic comparisons of cell values in each column by using a mechanism that is itself data-driven.  Because we use space partitioning algorithms to find surface forms that are potentially joinable, the search process is efficiently eliminates large parts of the vector space from consideration for any given queried vector.  Across all cell values that are considered for a join, the efficiency of the search is $\mathcal{O}(n\log{}n)$.

Specifically, the first part of our solution to the join problem involves building a deep learning model that performs metric learning.  For the closely related problem of face recognition and person re-identification, deep neural networks have been used successfully to build models that map faces of the same entity closer in vector space estimate relative to faces from other entities.  To test the generality of our approach, we tried to build a deep learning model model for names of people and companies. As in face recognition, our system for metric learning is built by training a `triplet siamese network' to learn to map different surface forms of the same entity closer in vector space than different forms of the same entity.  Entity names have very different properties from faces, so we needed to adapt the approaches used in face recognition to the problem of matching entity names with a novel method to choose the sample for training, and novel loss functions as objectives for the metric learning problem.  At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Douglas Noel Adams}) and a negative element (e.g., John Adams), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. A classification form of this function that simply learns whether two names belong to the same entity can be used to perform fuzzy joins, but then we would need a blocking strategy analogous to what has been used in the literature to identify the set of joinable pairs \cite{auto-join-joining-tables-leveraging-transformations}.

Our observation is that one can actually exploit what the siamese network produces to eliminate this blocking step altogether.  Specifically, the last hidden layer of the siamese network is effectively a `vector embedding' that contains the critical features needed for computing the distance estimate.  If such models are pre-built for a given entity type, we can then use the models to produce vector embeddings for all cell values in the two columns to be joined in a single linear pass, and index just the second column values with a space partitioning algorithm that lays out vectors in n-dimensional space.  Then, for each cell value in the first column to be joined, querying the k-nearest neighbors should provide the set of values in the second column that can be joined. In our work, we used approximate nearest neighbor algorithms, which have been applied successfully to billions of items \cite{JDH17}.

@@

Our key contributions are as follows:
\begin{itemize}
\item We outline a general approach to the problem of merging datasets for different surface forms of the same entity.  Our approach consists of building a deep neural network model for specific entity types and using a space partitioning algorithm to find the nearest neighbors to join.  To test the generality of our approach, we built triplet loss based siamese networks for ~200,000 people's names and ~70,000 company names respectively, drawn from Wikidata.  We show that such networks did learn to map surface forms of the same entity closer in vector space.  Specifically, if we 83\% and 79\% of variants of the entity name appeared in the top 20 neighbors for each entity.  Moreover, within the 20 neighbors, 77\% and 75\% of the positive names were closer to the anchor than negative names for people and companies respectively.  Our results suggest that deep learning models can be used successfully for data driven joins, assuming that such models are built for common entity types.
\item Training neural networks for deep metric learning is difficult because one cannot show quadratic combinations of the data to the system.  In face recognition, a number of techniques exist for triplet selection, and methods for selecting triples is an active area of research.  We developed a new technique for triplet selection for entity names because the techniques from face recognition do not generalize well to entity names.  We also developed a novel loss function to train the network to discriminate entity names from names of other entities.  We compare the results for this loss function with the ones that have been used for face recognition, and show that this loss function can perform much better than the existing functions for deep learning in this domain.
\item We also perform a complexity analysis of how such models can then be used to perform joins across different surface forms of the same entity efficiently, with a complexity of $\mathcal{O}(n\log{}n)$.  
\end{itemize}

The rest of the paper is organized as follows: Section~\ref{siamese networks} describes how we adapted siamese networks to the problem of entity matching, specifically the problem of triplet selection for entity names, and the need for a new type of loss function to improve metric learning for this problem, section ~\ref{joins} describes how to use these models for efficient joins across datasets, section ~\ref{dataset} describes characteristics of the datasets we used for training, as well as cleansing and data augmentation techniques we used to generate data, section ~\ref{results} provides the experimental results, section~\ref{related work} covers related work, and section~\ref{conclusions} describes future work and conclusions.

>>>>

In this paper, we propose a different approach to the problem of joining different surface representations of the same entity, inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a space partitioning algorithm to find the forms that can be potentially joined together.  The approach is general, in the sense that once a model has been built for a specific semantic type (e.g. people, companies or faces) it can be used for joining any two datasets which share that semantic type.  It is also efficient because it uses space partitioning algorithms (such as approximate nearest neighbor) to find surface forms that are potentially joinable, thus eliminating large parts of the vector space from consideration.  Further, implementations exist for nearest neighbor algorithms that have been applied to billions of vectors \cite{JDH17}, so the approach is practical for most datasets.

To test the feasability of these ideas, we used Wikidata as ground truth to build models for datasets with 1.1M people's names (about 200K identities) and 130K company names (70K identities).  The problem of mapping vectors for the same entity closer in vector space than vectors for other entities is known in the literature as deep metric learning.  Deep metric learning is known to be a difficult problem as studied in the space of face recognition and person-re-identification \cite{DBLP:conf/cvpr/SchroffKP15}.  As a result, there is a significant amount of research that has been targeted at two aspects of training these networks (a) how to choose samples for efficient learning, and (b) what constitutes a good loss function.  In building models for entity names, we developed a novel method for sample selection, and a novel loss function, which was effective in building these models.

As in face recognition, our system for metric learning is built by training a so-called `siamese triplet' network to learn to produce a small distance estimate for two surface forms for the same entity (positive pairs), and a large distance estimate for surface forms of different entities (negative pairs).  A key problem in effective training of such networks is the problem of how to select negative pair for training, because one cannot exhaustively show all negative pairs to the network.  In prior work for instance, this problem is solved by so-called `triplet mining' where negatives are gleaned after an all-pairs comparison of input vectors in a batch \cite{DBLP:conf/cvpr/SchroffKP15}.  Our novel approach to this problem is to re-use the approximate nearest neighbors algorithm to choose negatives for training.  Because our ultimate goal is to train the system to learn to map positives pairs closer and negative pairs further away from each other, we select negatives that appear in the nearest neighbor set for input vectors of a given entity, before any training has taken place.  This approach has two key benefits.  First, it lays out the entities based on their input vectors, and thus allows an efficient gathering of all nearest neighbors without a quadratic comparison of vectors to determine suitable negatives.  Second, it provides a baseline against which one can objectively measure the effects if any of training.  If for instance, the input vectors for surface forms of the same entity are already close together in vector space, little can be gained by a deep learning model.

We also investigated the effect of manipulating multiple loss functions which have been proposed in the literature for improving deep metric learning, e.g., the triplet loss \cite{DBLP:conf/cvpr/SchroffKP15}, improved loss \cite{Zhang:2016:DML:3088616.3088665}, and angular loss \cite{DBLP:journals/corr/abs-1708-01682} functions.  For the problem of building models for people's names, we found that a modification to the objective function proposed in \cite{DBLP:conf/cvpr/SchroffKP15} was better than three other functions that have been proposed in the literature.

@@

%Our key contributions are as follows:
%\begin{itemize}
%\item We outline a general approach to the problem of merging datasets for different surface forms of the same entity.  Our approach consists of building a deep neural network model for specific entity types and using a space partitioning algorithm to find the nearest neighbors to join.  To test the generality of our approach, we built triplet loss based siamese networks for 200K people's names and 70K company names respectively, drawn from Wikidata.  We show that such networks did learn to map surface forms of the same entity closer in vector space, with a recall rate of 81\% and 75\% after training for 20 neighbors.  Moreover, within the 20 neighbors, 93\% and 96\% of the positive names were closer to the anchor than negative names for people and companies respectively, suggesting good precision.  Our results suggest that deep learning models can be used successfully for data driven joins, assuming that such models are built for common entity types.
%\item Training neural networks for deep metric learning is difficult because one cannot show quadratic combinations of the data to the system.  In face recognition, a number of techniques exist for triplet selection, and methods for selecting triples is an active area of research.  We developed a new technique for triplet selection for entity names based on an approximate nearest neighbors algorithm over input vectors.  We also explored whether a modified loss function can be used to train the network more effectively, and found that it does in fact perform better for training on people's names.
%\item A key element of our argument is that once these embedding models have been built, they can be applied to perform joins efficiently for a new dataset.  We test if this is true with previously built face recognition models, because of the paucity of large datasets for companies and people, and the wealth of datasets in the face recognition literature.  For the most part, deep metric models for faces have been used with a classification layer have been used to identify a specific individual, or for clustering faces to measure goodness of the clustering algorithm.  Because we use these existing models for a slightly different purpose, we calculated the same metrics of precision and recall that we calculated for people and companies. We compared the performance of the Facenet model that was trained on VGGFace2 \cite{DBLP:conf/fgr/CaoSXPZ18} on the Labeled Faces in the Wild dataset \cite{Huang_labeledfaces}.  Our results show a recall of 80\%, and a precision of 93\%.  
%\end{itemize}

%The rest of the paper is organized as follows: Section~\ref{siamese networks} describes how we adapted siamese networks to the problem of entity matching, specifically the problem of triplet selection for entity names, and the need for a new type of loss function to improve metric learning for this problem, section ~\ref{joins} describes how to use these models for efficient joins across datasets, section ~\ref{dataset} describes characteristics of the datasets we used for training, as well as cleansing and data augmentation techniques we used to generate data, section ~\ref{results} provides the experimental results, section~\ref{related work} covers related work, and section~\ref{conclusions} describes future work and conclusions.

### 2018-08-28 paper2018/main.tex
draft of evaluation results
cleaning up

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\usepackage{subcaption}

@@

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Merging datasets through deep learning}

\author{Kavitha Srinivas \\ IBM Research
\And Yehuda Gale \\ Yeshiva University
\And Julian Dolby \\ IBM Research}

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is joining across columns that have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity to allow so-called `fuzzy' joins.  We propose a novel approach to performing merges inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a nearest neighbors algorithm to find the forms that can be potentially joined together.  Using Wikidata as our ground truth, we built models for datasets with approximately 1.1M people's names (200K identities) and 130K company names (70K identities), where the vector embeddings allow joins with precision of .91-.93 and recall of .70-.81. To build these models, we developed novel techniques for the selection of samples to train on, as well as a loss function, which we describe in the paper.  We also make the models available to the community for use in joining names or for aligning people or companies across multiple ontologies.

@@

\input{introduction2}
\input{siamese_networks}
\input{joins}
\input{datasets}

>>>>

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\usepackage{subcaption}
\usepackage{multirow}

@@

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Merging Datasets Through Deep learning}

%\author{Kavitha Srinivas \\ IBM Research
%\And Yehuda Gale \\ Yeshiva University
%\And Julian Dolby \\ IBM Research}

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is joining across columns that have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity.  We propose a novel approach to performing merges based on deep learning models.  Our approach depends on (a) creating a deep learning model that maps surface forms of an entity into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a nearest neighbors algorithm to find the forms that can be potentially joined together.  Using Wikidata as our ground truth, we built models for datasets with approximately 1.1M people's names (200K identities) and 130K company names (70K identities), where the vector embeddings allow for joins with precision of .91-.93 and recall of .70-.81. To build these models, we developed novel techniques for the selection of samples to train on, as well as a loss function, which we describe in the paper.  We make the models available to the community for aligning people or companies across multiple datasets.

@@

\input{introduction2}
\input{relatedWork}
\input{siamese_networks}
\input{joins}
\input{datasets}
\input{evaluation}

### 2018-08-28 paper2018/siamese_networks.tex
intro reworked
cleaning up

\section{Siamese networks}
\label{siamese networks}

Section~\ref{architecture} describes our overall network architecture for entity matching.  Section~\ref{loss_functions} enumerates the different loss functions which have been used in prior work, and a new loss function which appears to do better in metric learning for entities.  Section~\ref{triplet_selection} describes a novel triplet selection mechanisms for entity matching.

\subsection{Network Architecture}
\label{architecture}
\begin{figure}
\includegraphics[width=1.0\linewidth]{triplet_siamese_network}
\caption{Architecture of the siamese network}
\label{siamese_nets}
\end{figure}

Figure~\ref{siamese_nets} illustrates overall network architecture we used to build a system for fuzzy joins.  Input is initially triples of a name (anchor), another name for the same entity (positive) and a name of another entity (negative).  For the network, input vectors are computed for each entity assuming a maximum name length of 10 for each entity (i.e., no entity could have a name that spanned more than 10 tokens).  A 100 dimensional character embedding was computed for each token in the entity using pretrained embeddings \cite{hashimoto-jmt:2017:EMNLP2017}, which resulted in a 100 x 10 character encoding for each name used in a triplet.  These three input vectors from each triplet are fed to three identical networks that share the same weights.  Weight sharing ensures that the networks learn the same mapping function for any given input vector.  In our implementation, each of the three networks had 4 stacked layers of 128 unit Gated Recurrent Units (or GRUs) to capture the sequential nature of the input.

@@

The output of the last layer shown in the Figure~\ref{siamese_nets} as a dark gray layer produces an output vector embedding for the inputs.  These are fed to two layers which compute a euclidean distance between the \textit{anchor} and the \textit{positive} elements of a triplet (\textit{positive distance}), and the \textit{anchor} and the \textit{negative} elements of a triplet (\textit{negative distance}).  Conceptually, there are two objectives in metric learning, one to minimize \textit{positive distances}, and the other to maximize \textit{negative distances}.  As described below, this dual objective is achieved by different loss functions in the literature.

\subsection{Loss functions}
\label{loss_functions}

Let $\mathbf{x}$ represent an embedding for an entity name, and $\bf{x_{a}}$, $\bf{x_{p}}$, $\bf{x_{n}}$ reflect the vector embeddings of the \textit{anchor}, \textit{positive} and \textit{negative}, respectively.  We investigated four different loss functions to explore their effectiveness for the entity metric learning problem.

@@

\subsubsection{Triplet loss}

@@

\subsubsection{Improved Loss}

@@

\begin{equation}
  \psi_{triple} = \|\bf{x_{a}} - \bf{x_{p}}\|^2 - \hat{\alpha}
\label{psi}
\end{equation}
\begin{equation}
  \phi_{triple} = \|\bf{x_{a}} - \bf{x_{p}}\|^2 - (\|\bf{x_{a}}- \bf{x_{n}}\|^2 + \|\bf{x_{p}} - \bf{x_{n}}\|^2) / 2  + \alpha
\label{phi}
\end{equation}
\begin{equation}
  l_{triple} = max(0, \phi) + \lambda * max(0, \psi)
\label{improved_loss}
\end{equation}
The parameter $\hat{\alpha}$ for equation~\ref{psi} is set to 0.1, and $\lambda$ is set to .02 in equation~\ref{improved_loss}, and $\alpha$ was set to 1 in the paper \cite{Zhang:2016:DML:3088616.3088665}.  As in Schroff et al. \cite{DBLP:conf/cvpr/SchroffKP15}, the embeddings are normalized to 1.

\subsubsection{Angular Loss}

Wang et al. \cite{DBLP:journals/corr/abs-1708-01682} define a novel
angular loss function which is not based on pairwise distances, but
rather is based on the angles of the triangle formed by the
\textit{anchor}, \textit{positive} and the \textit{negative} triplet.
Conceptually, they point out that since the \textit{anchor} and the
\textit{positive} pairs belong to the same class, the angle formed by
the \textit{anchor}, \textit{negative}, and \textit{positive} elements
should be as small as possible within that triangle.  Their loss
function is an attempt to minimize this angle, as defined in the
equations below.  The rough idea is that moving the positive nearer
and the negative further away each reduce the angle, which we
illustrate in Figure~\ref{angular_loss} with the angle to shrink.
\begin{equation}
\bf{x_{c}} = (\bf{x_{a}} + \bf{x_{c}}) / 2
\end{equation}
\begin{equation}
l_{triplet} = \left[[\bf{x_{a}} - \bf{x_{p}}\|^2 - 4 \tan^2 \alpha \|\bf{x_{n}} - \bf{x_{c}}\|^2 \right])_{+}
\end{equation}

\subsubsection{Hyperbolic Tangent Loss}

>>>>

\section{Network Architecture}
\label{architecture}
\begin{figure}
\includegraphics[width=1.0\linewidth]{triplet_siamese_network}
\caption{Architecture of the siamese network}
\label{siamese_nets}
\end{figure}

Figure~\ref{siamese_nets} illustrates the siamese network architecture \cite{DBLP:conf/cvpr/SchroffKP15} we used to build a system for merging data.  Input is initially triples of a name (\textit{anchor}), another name for the same entity (\textit{positive}) and a name of another entity (\textit{negative}).  For the network, input vectors are computed for each entity assuming a maximum name length of 10 for each entity (i.e., no entity could have a name that spanned more than 10 tokens).  A 100 dimensional character embedding was computed for each token in the entity using pretrained character embeddings \cite{hashimoto-jmt:2017:EMNLP2017}, which resulted in a 100 x 10 character encoding for each name used in a triplet.  These three input vectors from each triplet are fed to three identical networks that share the same weights.  Weight sharing ensures that the networks learn the same mapping function for any given input vector.  In our implementation, each of the three networks had 4 stacked layers of 128 unit Gated Recurrent Units (or GRUs) to capture the sequential nature of the input.

@@

The output of the last layer shown in the Figure~\ref{siamese_nets} as a dark gray layer produces an output vector embedding for the inputs.  These are fed to two layers which compute a euclidean distance between the \textit{anchor} and the \textit{positive} elements of a triplet (\textit{positive distance}), and the \textit{anchor} and the \textit{negative} elements of a triplet (\textit{negative distance}).  Conceptually, there are two objectives in metric learning, one to minimize \textit{positive distances}, and the other to maximize \textit{negative distances}.  As described below, this dual objective can be achieved by different loss functions.

\section{Loss functions}
\label{loss_functions}

Let $\mathbf{x}$ represent an embedding for an entity name, and $\bf{x_{a}}$, $\bf{x_{p}}$, $\bf{x_{n}}$ reflect the vector embeddings of the \textit{anchor}, \textit{positive} and \textit{negative}, respectively.  We investigated four different loss functions, three of which have been used in prior literature to explore their effectiveness for the entity metric learning problem.

@@

\subsection{Triplet loss}

@@

\subsection{Improved Loss}

@@

\begin{equation}
  \psi_{triple} = \|\bf{x_{a}} - \bf{x_{p}}\|^2 - \hat{\alpha}
\label{psi}
\end{equation}
\begin{equation}
  \phi_{triple} = \|\bf{x_{a}} - \bf{x_{p}}\|^2 - (\|\bf{x_{a}}- \bf{x_{n}}\|^2 + \|\bf{x_{p}} - \bf{x_{n}}\|^2) / 2  + \alpha
\label{phi}
\end{equation}
\begin{equation}
  l_{triple} = max(0, \phi) + \lambda * max(0, \psi)
\label{improved_loss}
\end{equation}
The parameter $\hat{\alpha}$ for equation~\ref{psi} is set to 0.1, and $\lambda$ is set to .02 in equation~\ref{improved_loss}, and $\alpha$ was set to 1 in the paper \cite{Zhang:2016:DML:3088616.3088665}.  As in Schroff et al. \cite{DBLP:conf/cvpr/SchroffKP15}, the embeddings are normalized to 1 although the actual paper does not make it clear.

\subsection{Angular Loss}

Wang et al. \cite{DBLP:journals/corr/abs-1708-01682} define a novel
angular loss function which is not based on pairwise distances, but
rather is based on the angles of the triangle formed by the
\textit{anchor}, \textit{positive} and the \textit{negative} triplet.
Conceptually, they point out that since the \textit{anchor} and the
\textit{positive} pairs belong to the same class, the angle formed by
the \textit{anchor}, \textit{negative}, and \textit{positive} elements
should be as small as possible within that triangle.  Their loss
function is an attempt to minimize this angle, as defined in the
equations below.  The rough idea is that moving the positive nearer
and the negative further away each reduce the angle, which we
illustrate in Figure~\ref{angular_loss} with the angle to shrink.
\begin{equation}
\bf{x_{c}} = (\bf{x_{a}} + \bf{x_{p}}) / 2
\end{equation}
\begin{equation}
l_{triplet} = \left[[\bf{x_{a}} - \bf{x_{p}}\|^2 - 4 \tan^2 \alpha \|\bf{x_{n}} - \bf{x_{c}}\|^2 \right])_{+}
\end{equation}

\subsection{Hyperbolic Tangent Loss}

### 2018-08-28 paper2018/triplet_selection.tex
cleaning up

\subsection{Triplet Selection}
Deep metric learning for faces is a difficult problem for neural networks to learn because improper sample selection can make it difficult for the network to converge.  We describe a popular approach from the face recognition literature first before we describe a new mechanism for triplet selection.

\subsubsection{Batch Based Triple Selection}
\begin{figure}
\includegraphics[width=1.0\linewidth]{triplet_selection}
\caption{Batch based triplet selection}
\label{triplet_selection}
\end{figure}

@@

\subsubsection{Metric Based Triplet Selection}
Although the computation of there is $argmax^i \| x_{a} - x_{p}^i \|^2$ and $argmin^i \| x_{a} - x{_n}^i \|^2$ is infeasible because it involves quadratic comparisons within a dataset, we observe that the use of approximate nearest neighbor (ANN) indexes can actually make this problem tractable for the selection of triples.

@@

Assuming one has the entire dataset indexed in an ANN index, the problem of triplet selection can be simplified by asking the ANN index for the top k-nearest neighbors of an \textit{anchor}, where k is given by the number of triplets that one desires to generate for each anchor.  As in earlier work, selection of \textit{hard positives} is not really relevant because all positive data should be used to teach the network the right function.  Selection of negatives is the set of all nearest neighbors that are \textit{negatives}.  There is no explicit attempt to filter out\textit{hard negatives} in the approach.  The overall idea is that teh set of negatives that appear in the nearest neighbor set at input are in fact the most important elements for the network to learn to discriminate from positives.  Focusing on these elements, regardless of whether they are \textit{hard} or \textit{semi-hard} should lead to better learning.

An ANN-based strategy works better than the batch based triplet selection approach on two counts.  First, the entire dataset is in fact used in the triplet selection process.  Second, for certain types of datasets, the batch based triplet selection mechanism will actually fail if it is focused on finding \textit{semi-hard negatives}.  In particular, anytime \textit{positive distances} are on average much greater than \textit{negative} distances, there are very few \textit{semi-hard negatives} because most positives are in fact more distant than the negatives.  We observed this property in the people's name matching dataset.  If we take the character embeddings of each of the names in our name matching dataset, index it in an ANN, and then ask for the top-k neighbors we find somewhat troubling properties for an approach that tries to select \textit{semi-hard} negatives.  First, \textit{positives} appeared in the top 20 neighbors only in 3.2\% of the dataset.  Explicitly computing the distance between the \textit{anchor} and the \textit{positive} yielded a mean of 9.07, with a standard deviation of 3.07.  The mean distances between the \textit{anchor} and the \textit{negatives} for those negatives in the top 20 was 3.16 with a standard deviation of 1.17.  Given that only 3.2\% of positives actually even appeared in the top-K nearest neighbors, selection of \textit{semi-hard} negatives is reduced to a very small fraction of the dataset.  A k-nearest neighbor approach to triplet selection on the other hand works on this difficult dataset as well, by picking all positive triplets, and a set of \textit{negatives} based on the top-k nearest neighbors.  As we show later this approach generalizes to datasets such as company names where such a strong reversal of \textit{positive} and \textit{negative} distances does not exist.

>>>>

\section{Triplet Selection}
Deep metric learning for faces is a difficult problem for neural networks to learn because improper sample selection can make it difficult for the network to converge.  We describe a popular approach from the face recognition literature first before we describe a new mechanism for triplet selection.

\subsection{Batch Based Triple Selection}
\begin{figure}
\includegraphics[width=1.0\linewidth]{triplet_selection}
\caption{Batch based triplet selection}
\label{triplet_selection}
\end{figure}

@@

\subsection{Metric Based Triplet Selection}
Although the computation of there is $argmax^i \| x_{a} - x_{p}^i \|^2$ and $argmin^i \| x_{a} - x{_n}^i \|^2$ is infeasible because it involves quadratic comparisons within a dataset, we observe that the use of approximate nearest neighbor (ANN) indexes can actually make this problem tractable for the selection of triples.

@@

Assuming one has the entire dataset indexed in an ANN index, the problem of triplet selection can be simplified by asking the ANN index for the top k-nearest neighbors of an \textit{anchor}, where k is given by the number of triplets that one desires to generate for each anchor.  As in earlier work, selection of \textit{hard positives} is not really relevant because all positive data should be used to teach the network the right function.  Selection of negatives is the set of all nearest neighbors that are \textit{negatives}.  There is no explicit attempt to filter out\textit{hard negatives} in the approach.  The overall idea is that the set of negatives that appear in the nearest neighbor set at input are in fact the most important elements for the network to learn to discriminate from positives.  Focusing on these elements, regardless of whether they are \textit{hard} or \textit{semi-hard} should lead to better learning.

An ANN-based strategy works better than the batch based triplet selection approach on two counts.  First, the entire dataset is in fact used in the triplet selection process.  Second, it yields an important baseline to assess what if any learning was performed by the neural network in mapping input vectors to a different space.
